---
title: "Activity 3"
author: "Name here"
output: pdf_document
---

### Q1.
Suppose we have a probability mass function for weekly winter snowfall at bridger bowl that temperature in Hyalite that is Poisson with mean of 12 inches. This 

```{r}
library(tidyverse)
snow_seq <- 0:50
tibble(vals = c(dpois(snow_seq, 12)),
       `Weekly snowfall (inches)` = snow_seq) %>%
  ggplot(aes(x = `Weekly snowfall (inches)`, y = vals)) +
  geom_col() + theme_bw() +
  ggtitle(expression(paste('Distribution for snowfall: Poisson(12)')))
```

Answer the following questions with an numeric answer and a description of why that answer is true.

- According to this distribution, what is the probability of a week having 12 inches of snow? (Hint `dpois()`)

*The probability is `r round(dpois(12,12),2)`. With a discrete variable, this is the height of the probability mass function. We can use `dpois` in R.*

- According to this distribution, what is the probability of a week having more than 12 inches of snow? (Hint `ppois()`)

*The probability is `r round(1 - ppois(12,12),2)`. Here we are finding the area of the mass function which is greater than 12. We can use `ppois` in R.*

- According to this distribution, what is the probability of a week having 33 inches of snow (As of Jan 29, Bridger Bowl has reported 33 inches in last week)?

*The probability of this occurrence is very small, specifically it is `r dpois(33,12)`.*

- Based on the last question, do you have any concerns with this function for snowfall?

*The probability of this occurrence is very small, specifically it is `r dpois(33,12)`. This suggests either the last week was really unusual or the probability model has some issues. Based on my experiences, it is likely a little of both. However, this last week wasn't as unusual as suggested by the model; rather, a model that has greater variability (maybe negative binomial) might better explain nature.*

### Q2.
Suppose we have a probability distribution for average temperature in Hyalite that is Normal with mean = 10 and standard deviation = 7.

```{r}
temp_seq <- seq(-40,40, length.out = 100)
tibble(vals = c(dnorm(temp_seq, 10, 7)),
       temperature = temp_seq) %>%
  ggplot(aes(x = temperature, y = vals)) +
  geom_line() + theme_bw() +
  ggtitle(expression(paste('Distribution for mean temperature: N(10,',7^2,')')))
```

Answer the following questions with an numeric answer and a description of why that answer is true.

- What is the probability that the temperature is greater than 10 degrees?

*The probability is $\frac{1}{2}$. The normal distribution is symmetric, so we have equal probability of a value being greater and lower than the mean.*

- What is the probability that the temperature is less than 0 degrees? (Hint `pnorm`)

*The probability is `r round(pnorm(0,10,7),2)`. We are actually integrating the area of the curve less than 0 degrees. This integration can be done in R using `pnorm()`.*


### Q3.

Yahtzee is a dice game where players roll 5 dice. A yahtzee (all 5 dice with the same value) is worth 50 points and a large straight (a run of 5 consecutive values) is worth 40 points.  Use a Monte Carlo technique to calculate the probability of each of these outcomes, given a single roll.

```{r}
num_sims <- 1000000

Yahtzee <- straight <- rep(FALSE, num_sims)

for (i in 1:num_sims){
  roll <- sample(6, 5, replace = T)
  Yahtzee[i] = length(unique(roll)) == 1
  straight[i] = (length(unique(roll)) == 5 & !(1 %in% roll)) | (length(unique(roll)) == 5 & !(6 %in% roll))
}

mean(Yahtzee)
mean(straight)
```

*The probability of a Yahtzee is `r round(mean(Yahtzee),5)` and the probability of a straight is `r round(mean(straight),5)`.*
