---
title: "Activity 4"
format: gfm
editor: source
editor_options: 
  chunk_output_type: console
---

### Week 3 Recap

-   Thinking with distributions (intervals, mean, dispersion)

-   Monte Carlo Techniques

-   Probability mass functions and probability density functions

------------------------------------------------------------------------

### Week 4 Overview: Conditional Probability and Bayes Rule

-   Marginal, Joint, Conditional Probability

-   Bayes Rule

-   Bayes Rule with data

------------------------------------------------------------------------


The highest density interval (HDI) corresponds to the shortest possible interval, for a specified probability level. What would be the 95% HDI for the following discrete distributions? Note with discrete distributions you can calculate these exactly, but we can also use `hdi` from `HDInterval` for approximate, Monte Carlo - based intervals

#### 1a. Poisson(4)

```{r, message = F}
library(tidyverse)
library(knitr)
x_seq <- 0:20
pois4 <- tibble(x = x_seq, prob = dpois(x_seq,4)) 

pois4 %>%
  ggplot(aes(x=x,y=prob)) + geom_col() + theme_bw()

pois4 |>
  kable()

pois4 |>
  filter(x > 0, x <= 8) |>
  summarize(sum(prob))

```

This would be [1, 8]

```{r}
library(HDInterval)

hdi(rpois(100000,4))

```


#### 1b. NB(4, 100)

```{r}
x_seq <- 0:20
nb4 <- tibble(x = x_seq, prob = dnbinom(x_seq,mu =4, size = 100)) 

nb4 %>%
  ggplot(aes(x=x,y=prob)) + geom_col() + theme_bw()

nb4 |>
  kable()

nb4 |>
  filter(x > 0, x <= 8) |>
  summarize(sum(prob))
```

This would be approximately [1,8]


```{r}
hdi(rnbinom(100000, mu =4, size = 100))
```

#### 1c. NB(4, 1)

```{r}
x_seq <- 0:20
nb4_1 <- tibble(x = x_seq, prob = dnbinom(x_seq,mu =4, size = 1)) 
nb4_1 %>%
  ggplot(aes(x=x,y=prob)) + geom_col() + theme_bw()

nb4_1 |> kable()

nb4_1 |>
  filter(x <= 13) |>
  summarize(sum(prob))
```


This would be approximately [0, 13]

```{r}
hdi(rnbinom(100000, mu =4, size = 1))
```

#### 1d. Pois(6)

```{r}
x_seq <- 0:20
pois6 <- tibble(x = x_seq, prob = dpois(x_seq,6)) 

pois6 %>%
  ggplot(aes(x=x,y=prob)) + geom_col() + theme_bw()

pois6 |>
  kable()

pois6 |>
  filter(x >=2 , x <= 11) |>
  summarize(sum(prob))

```

This would be [2, 11]

```{r}
library(HDInterval)

hdi(rpois(100000,6))

```

---

Use a dataset containing homes in the Seattle, WA area [http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv](http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv) for this question.

```{r}
seattle <- read_csv('http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv') %>%
  mutate(bath_category = case_when(
    bathrooms <= 2 ~ '0 - 2',
    bathrooms > 2 & bathrooms <=4 ~ '2 - 4',
    bathrooms > 4 ~ 'more than 4'),
    bed_category = case_when(
    bedrooms == 0 ~ '0',
    bedrooms > 0 & bedrooms <= 2 ~ '1-2',
    bedrooms > 2 & bedrooms <= 4 ~ '3-4',
    bedrooms > 4 ~ 'more than 5'
   )
  )
```


Use the two-by-two table containing bathrooms (grouped as: 0 - 2, more than 2 - 4, more than 4) and bedrooms (0, 1-2, 3-4, 5 or more).
```{r}
table(seattle$bath_category, seattle$bed_category) 
```


```{r}
round(table(seattle$bath_category, seattle$bed_category) / nrow(seattle),3)
```


#### 2a.
Compute marginal probability of having 0 - 2 bathrooms

The probability is `r .003 + .148 + .382 + .015`

#### 2b. 
Compute joint probability of having 0 - 2 bathroom and  3-4 bedrooms

The probability is .382

#### 2c. 
Compute conditional probability of having 3-4 bedrooms given that it has 0 - 2 bathrooms

The probability is `r round(.382/ (.003 + .148 + .382 + .015),3)`


#### 2d. 
Are bathrooms and bedrooms independent? Why or why not.

*No the probability of the number of bedrooms depends on the number of bathrooms. In particular there are different values for parts a and c.*


---

(DBDA 4.5B)

Use a normal curve to describe the following belief. Suppose you believe that women's heights follow a bell-shaped distribution, centered at 162 cm with about 2/3rds of all women having heights between 147 and 177. What should be the $\mu$ and $\sigma$ values?

- $\mu$ = 162
- $\sigma$ = 15 (as about 2/3rd of observations fall within 1 sd of the mean)

---

Assume your roommate has taken, and tested positive, for an at home antigen Covid test. Let's assume the test was from Cochrane (data)[https://www.cochrane.org/CD013705/INFECTN_how-accurate-are-rapid-antigen-tests-diagnosing-covid-19] with the following properties:

- P[Test = + | Disease = +] = .82
- PP[Test = - | Disease = -] = .995

#### 4a.

If the overall population prevalence was 5%, what is the probability that your roommate has COVID-19?


```{r}
p.t.plus <- .82
p.t.minus <- 1 - .995
p.theta <- .05
p.theta.true <- p.t.plus * p.theta / (p.t.plus * p.theta + p.t.minus * (1 - p.theta))
```
The probability would be about `r round(p.theta.true,3)`.

#### 4b.

Now assume that your other roommate took a test (with the same characteristics) and that test was negative. What is the probability the other roommate has COVID-19?


```{r}
p.t.plus <- .82
p.t.minus <- 1 - .995
p.theta <- .05
p.theta.neg <- (1 - p.t.plus) * (p.theta) / ((1 - p.t.minus) * (1- p.theta) + (1 - p.t.plus) * (p.theta))
```

The probability would be about `r round(p.theta.neg,3)`