---
title: "Week 12 Activity: Regression"
output: pdf_document
---

```{r setup, include=FALSE}
library(rjags)
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

```

On Thursday we will use historical data to predict NCAA basketball games. For this activity we will generate (simulate) synthetic data.


### 1. 

Simulate synthetic data that represents historical NCAA games. In particular, let's consider the model

$$pointdiff = \beta_0 + \beta_1 x_{seeddiff} + \epsilon; \epsilon \sim N(0, \sigma^2)$$

where:

- $\beta_0$ = 0
- $\beta_1$ = 2
- $\sigma = 12$


```{r}
set.seed(04032023)
num_games <- 960 # approximately 15 years
seeddiff <- rep(0:15, each = 60)


beta1 <- 2
sigma <- 12

```


### 2.

Interpret the three coefficients in the model specified in part 1.

- $\beta_0$ = 
- $\beta_1$ = 
- $\sigma$ = 

### 3. 

Create a visualization of your point differential versus seed differential.


### 4. 

Specify prior distributions on your parameters.

- $\beta_0 \sim$ 
- $\beta_1 \sim$ 
- $\sigma \sim$ 

### 5. 

Write JAGS code to fit this model. Output the results.

### 6.

Summarize your findings. Write a paragraph so that a college basketball coach could understand.


### 7.

Now we will construct a posterior predictive distribution. This will enable us answers questions like "What is the probability that a one seed is upset by a 16 seed (15 point difference in seeds.) In particular, construct a posterior predictive distribution, conditional on the following scenarios:

- Seed difference = 15
- Seed difference = 7 (most commonly 1 vs. 8)
- Seed difference = 1

Use those distributions to compute the probability of an upset occuring.

