---
title: "Activity 4"
author: "Name here"
output: pdf_document
---

### Q1. 
Describe the differences between probability density functions and probability mass functions.

### Q2. 
The highest density interval (HDI) corresponds to the shortest possible interval, for a specified probability level. What would be the 95% HDI for the following discrete distributions?

#### a. 
Poisson(4)
```{r, message = F}
library(tidyverse)
x_seq <- 0:20
tibble(x = x_seq, prob = dpois(x_seq,4)) %>%
  ggplot(aes(x=x,y=prob)) + geom_col() + theme_bw()
```


#### b. 
NB(4, 100)

```{r}
x_seq <- 0:20
tibble(x = x_seq, prob = dnbinom(x_seq,mu =4, size = 100)) %>%
  ggplot(aes(x=x,y=prob)) + geom_col() + theme_bw()
```


#### d. 
NB(4, 1)

```{r}
x_seq <- 0:20
tibble(x = x_seq, prob = dnbinom(x_seq,mu =4, size = 2)) %>%
  ggplot(aes(x=x,y=prob)) + geom_col() + theme_bw()
```



#### 2. 
Pois(6)

```{r}
tibble(x = x_seq, prob = dpois(x_seq,6)) %>%
  ggplot(aes(x=x,y=prob)) + geom_col() + theme_bw()
```

This would be approximately (2, 11) (`sum(dpois(2:11,6))`)


### Q3.

Use a dataset containing homes in the Seattle, WA area [http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv](http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv) for this question.

```{r}
seattle <- read_csv('http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv') %>%
  mutate(bath_category = case_when(
    bathrooms <= 2 ~ '0 - 2',
    bathrooms > 2 & bathrooms <=4 ~ '2 - 4',
    bathrooms > 4 ~ 'more than 4'),
    bed_category = case_when(
    bedrooms == 0 ~ '0',
    bedrooms > 0 & bedrooms <= 2 ~ '1-2',
    bedrooms > 2 & bedrooms <= 4 ~ '3-4',
    bedrooms > 4 ~ 'more than 5'
   )
  )
```


Use the two-by-two table containing bathrooms (grouped as: 0 - 2, more than 2 - 4, more than 4) and bedrooms (0, 1-2, 3-4, 5 or more).
```{r}
table(seattle$bath_category, seattle$bed_category) 
```



#### a.
Compute marginal probability of having 0 - 2 bathrooms


#### b. 
Compute joint probability of having 0 - 2 bathroom and  3-4 bedrooms


#### c. 
Compute conditional probability of having 3-4 bedrooms given that it has 0 - 2 bathrooms



#### d. 
Are bathrooms and bedrooms independent? Why or why not.


### Q4. (DBDA 4.5B)

Use a normal curve to describe the following belief. Suppose you believe that women's heights follow a bell-shaped distribution, centered at 162 cm with about 2/3rds of all women having heights between 147 and 177. What should be the $\mu$ and $\sigma$ values?


### Q5.

Assume your roommate has taken, and tested positive, for an at home antigen Covid test. Let's assume the test was from Cochrane (data)[https://www.cochrane.org/CD013705/INFECTN_how-accurate-are-rapid-antigen-tests-diagnosing-covid-19] with the following properties:

- P[Test = + | Disease = +] = .82
- PP[Test = - | Disease = -] = .995

#### a.

If the overall population prevalence was 5%, what is the probability that your roommate has COVID-19?

#### b.

Now assume that your other roommate took a test (with the same characteristics) and that test was negative. What is the probability the other roommate has COVID-19?

