---
title: "Activity 7: key"
format: gfm
editor: source
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: FALSE
library(tidyverse)
library(rjags)

```


### Week 6 Recap: Bayesian Analysis with MCMC

-   jags

-   Bayesian Analysis with MCMC


------------------------------------------------------------------------

### Week 7 Overview: Normal Distribution

-   Bayes Rule with data

-   "Calculus"

-   Bayesian Data Analysis with binary data

------------------------------------------------------------------------

```{r, include =F}
library(tidyverse)
library(rjags)
```

This analysis will focus on small dataset containing information from Indeed.com, which can be accessed using [http://www.math.montana.edu/ahoegh/teaching/stat491/data/bzn_jobs.csv](http://www.math.montana.edu/ahoegh/teaching/stat491/data/bzn_jobs.csv).

```{r}
bzn_jobs <- read_csv('http://www.math.montana.edu/ahoegh/teaching/stat491/data/bzn_jobs.csv')
```

This dataset contains the following variables:

    - jobAgeDays: number of days the job has been posted on Indeed.com
    - normTitle: name of job position (registered nurse, sales associate, truck driver)
    - estimatedSalary: estimated annual salary
    - localClicks: number of people clicking on job posting

## Bayesian ANOVA
For this question we will fit a regression analysis (ANOVA) to model estimated salary across three diffferent job types. variables as predictors.

#### a. Data Viz

Create a figure of salary by `normTitle`. It is good practice to show all data points.


#### b.
Interpret the following R output.

```{r}
anova_fit <- lm(estimatedSalary ~ normTitle - 1, data = bzn_jobs)
summary(anova_fit)
confint(anova_fit)
```


#### c. 

Select and Justify a sampling model for your response.



#### d. 

Explain the purpose of this model - you can assume you talking to a freshman in high school.


#### e. 
State and Justify Priors Used for your  Model



#### e. 

Finish this JAGS code to fit the Posterior Distribution for this Model and print the results

```{r}
model_anova<- "model{
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(beta1 * x1[i] + beta2 * x2[i] + beta3 * x3[i], 1/sigma^2) 
  }
  
  # Prior
  beta1 ~ 
  beta2 ~ 
  beta3 ~ 
  sigma ~ 
}"
```


#### f. 
Visualize your results in some fashion - this can look similar to your figure in part a.


#### g.

Compare your interval results, in part f, with those from part b


#### h.

Explain the results of this model - you can assume you talking to a freshman in high school.

