---
title: "Activity 7"
author: "Name here"
output: pdf_document
---

```{r, include =F}
library(tidyverse)
library(rjags)
```

This analysis will focus on small dataset containing information from Indeed.com, which can be accessed using [http://www.math.montana.edu/ahoegh/teaching/stat491/data/bzn_jobs.csv](http://www.math.montana.edu/ahoegh/teaching/stat491/data/bzn_jobs.csv).

```{r}
bzn_jobs <- read_csv('http://www.math.montana.edu/ahoegh/teaching/stat491/data/bzn_jobs.csv')
```

This dataset contains the following variables:

    - jobAgeDays: number of days the job has been posted on Indeed.com
    - normTitle: name of job position (registered nurse, sales associate, truck driver)
    - estimatedSalary: estimated annual salary
    - localClicks: number of people clicking on job posting

## Bayesian ANOVA
For this question we will fit a regression analysis (ANOVA) to model estimated salary across three different job types.

#### a. Data Viz

Create a figure of salary by `normTitle`. It is good practice to show all data points.


#### b.
Interpret the following R output.

```{r}
anova_fit <- lm(estimatedSalary ~ normTitle - 1, data = bzn_jobs)
summary(anova_fit)
confint(anova_fit)
```

#### c. 

Select and Justify a sampling model for your response.


#### d. 

Explain the purpose of this model - you can assume you talking to a freshman in high school.

#### e. 
State and Justify Priors Used for your  Model


#### e. 

Modify existing JAGS code to fit this model.

```{r}
indicator_data <- model.matrix(estimatedSalary~normTitle - 1, data = bzn_jobs)
```


```{r}
model_anova<- "model{
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(beta1 * x1[i] + beta2 * x2[i] + beta3 * x3[i], 1/sigma^2) 
  }
  
}"
```


#### f. (4 points) 
Using your JAGS code to fit the Posterior Distribution for this Model and print the results


#### g. 
Visualize your results in some fashion.


#### h.

Compare your interval results, in part f, with those from part b


#### i.

Explain the results of this model - you can assume you talking to a freshman in high school.

