---
title: "Week 12 Activity: Regression"
output: pdf_document
---

```{r setup, include=FALSE}
library(rjags)
library(tidyverse)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

```

On Thursday we will use historical data to predict NCAA basketball games. For this activity we will generate (simulate) synthetic data.


### 1. 

Simulate synthetic data that represents historical NCAA games. In particular, let's consider the model

$$pointdiff = \beta_0 + \beta_1 x_{seeddiff} + \epsilon; \epsilon \sim N(0, \sigma^2)$$

where:

- $\beta_0$ = 0
- $\beta_1$ = 2
- $\sigma = 12$


```{r}
set.seed(04032023)
num_games <- 960 # approximately 15 years
seeddiff <- rep(0:15, each = 60)


beta1 <- 2
sigma <- 12

pointdiff <- rnorm(num_games, mean = beta1 * seeddiff, sd = sigma)

data_out <- tibble(seeddiff = seeddiff,
                   pointdiff = pointdiff) 

```


### 2.

Interpret the three coefficients in the model specified in part 1.

- $\beta_0$ = expected point differential for teams with the same seed
- $\beta_1$ = multiplicative coefficient for point differential as a function of seed difference
- $\sigma = 12$ = standard deviation in point differential. Most games would be $\pm$ 2 standard deviations (or 24 points) from the mean.

### 3. 

Create a visualization of your point differential versus seed differential.
```{r}
data_out %>% ggplot(aes(y = pointdiff, x = seeddiff), method = 'loess', formula = 'y ~ x') + geom_point() + geom_smooth() + theme_bw()

```


### 4. 

Specify prior distributions on your parameters.

- $\beta_0 \sim$ Normal(0,$.001^2$)
- $\beta_1 \sim$ Normal(1, $3^2$)
- $\sigma \sim$ Unif(0, 100)

### 5. 

Write JAGS code to fit this model. Output the results.

```{r}
#Prior parameters
M0 <- 0
S0 <- .001
M1 <- 1
S1 <- 3
C <- 100

# Store data
dataList = list(y = data_out$pointdiff, 
                x = data_out$seeddiff,
                N = nrow(data_out), 
                M0 = M0, S0 = S0,
                M1 = M1, S1 = S1, C = C)

# Model String
modelString = "model {
  for ( i in 1:N ) {
    y[i] ~ dnorm(beta0 + beta1 * x[i], 1/sigma^2) # sampling model
  }
  beta0 ~ dnorm(M0,1/S0^2)
  beta1 ~ dnorm(M1, 1 / S1^2)
  sigma ~ dunif(0,C)
} "
writeLines( modelString, con='NORMmodel.txt')


# Runs JAGS Model
jags.norm <- jags.model( file = "NORMmodel.txt", data = dataList,
                         n.chains = 2, n.adapt = 1000)
update(jags.norm, n.iter = 1000)

coda.norm <- coda.samples( jags.norm, variable.names = c('beta0','beta1', 'sigma'), n.iter = 5000)

summary(coda.norm)
```

### 6.

Summarize your findings. Write a paragraph so that a college basketball coach could understand.

**For each unit difference in seed the expected point differential would increase by about 2 points. For teams of the same seed, the expected point differential would be 0 points. As you know with the NCAA tournament there is the potential for upsets. This is illustrated by the $\sigma$ term in our model. For a given point spread, it would not be unlikely to see differences of about 20 points, either way, from the expected point spread.

### 7.

Now we will construct a posterior predictive distribution. This will enable us answers questions like "What is the probability that a one seed is upset by a 16 seed (15 point difference in seeds.) In particular, construct a posterior predictive distribution, conditional on the following scenarios:

- Seed difference = 15
- Seed difference = 7 (most commonly 1 vs. 8)
- Seed difference = 1

Use those distributions to compute the probability of an upset occuring.

```{r}
#Prior parameters
M0 <- 0
S0 <- .001
M1 <- 1
S1 <- 3
C <- 100

# Store data
dataList = list(y = data_out$pointdiff, 
                x = data_out$seeddiff,
                N = nrow(data_out), 
                M0 = M0, S0 = S0,
                M1 = M1, S1 = S1, C = C)

# Model String
modelString_pp = "model {
  for ( i in 1:N ) {
    y[i] ~ dnorm(beta0 + beta1 * x[i], 1/sigma^2) # sampling model
    pp[i] ~ dnorm(beta0 + beta1 * x[i], 1/sigma^2)
  }
  pp1 ~ dnorm(beta0 + beta1, 1/sigma^2)
  pp7 ~ dnorm(beta0 + beta1 * 7, 1/sigma^2)
  pp15 ~ dnorm(beta0 + beta1 * 15, 1/sigma^2)
  beta0 ~ dnorm(M0,1/S0^2)
  beta1 ~ dnorm(M1, 1 / S1^2)
  sigma ~ dunif(0,C)
} "
writeLines( modelString_pp, con='NORMmodel.txt')


# Runs JAGS Model
jags.norm <- jags.model( file = "NORMmodel.txt", data = dataList,
                         n.chains = 2, n.adapt = 1000)
update(jags.norm, n.iter = 1000)

coda.norm <- coda.samples( jags.norm, variable.names = c('beta0','beta1', 'sigma','pp1','pp7','pp15','pp'), n.iter = 5000)

pp_out <- tibble(vals = c(coda.norm[[1]][,'pp1'],
                          coda.norm[[1]][,'pp7'],
                          coda.norm[[1]][,'pp15']),
                 type = rep(c('seed diff 1', 
                              'seed diff 7', 
                              'seed diff 15'),
                            each = 5000))

pp_out %>% group_by(type) %>% 
  summarize(`upset prob` = mean(vals < 0)) %>%
  kable(digits = 3)

pp_out %>% ggplot(aes(x = vals)) +
  geom_histogram() + facet_grid(type~.)
```

