---
title: "Activity 14"
format: gfm
editor: source
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: FALSE
library(tidyverse)
library(rjags)
library(gridExtra)
```


### Last-Week's Recap

- GLMs
- Posterior predictive distributions


### This week

- Hierarchical Models

### Next week

- Project Analysis (for next Tuesday - no videos)
- Final Exam Review (Tuesday)
- In class exam (Thursday)
- Take Home Exam 

---

```{r setup, include=FALSE}
library(knitr)
library(rjags)
library(runjags)
library(tidyverse)
library(arm)
set.seed(04262023)
knitr::opts_chunk$set(echo = TRUE)
```


Similar to the Space Jam themed exercises previously For this activity we will use NBA free throw shooting data to compare a few different approaches for analyzing this dataset.

```{r, echo = F}
playoff.free.throws <- tibble(
  player = c('James','Curry','Irving','Durant','Wall','Leonard','Beal','Harden','Love','Horford','Aldridge','Green','Paul','Olynyk','Parker','Favors','Jordan','Oladipo'),
  position = c('F','G','G','F','G','G','G','G','F','F','F','F','G','F','G','F','F','G'),
  FTA = c(162,114,84,103,93,102,61,115,75,29,55,67,33,30,14,23,56,6),
  FTM = c(113,103,76,92,78,95,50,101,63,22,42,46,29,22,14,11,22,6)) %>% 
  arrange(position, player) |>
  filter(position == 'G')
playoff.free.throws %>% kable(digits = 2)
```

---

1. Fit a frequentist model to estimate free throw shooting percentage, $\theta,$ for the players Curry, Beal, and Oladipo. Write the sampling model you've assumed with your code (hint: `glm()` is one possible route). Report a point estimate and uncertainty intervals for each player.
\vfill


---

2. Specify independent models using a uniform prior for $\theta$ for the players Curry, Beal, and Oladipo. Write the the sampling model and priors for these models. Recall you can find the posterior here analytically without using MCMC. Find posterior means and HDI intervals for this approach. Recall the HPDinterval function can be applied directly to samples from a beta distribution as `HPDinterval(mcmc(data  = rbeta(n = 5000, shape1 = a.star, shape2 = b.star)))`

---

3. Specify independent models using an informative prior, of your choice, for $\theta$ for the players Curry, Beal, and Oladipo. Write the the sampling model and priors for these models. Recall you can find the posterior here analytically without using MCMC. Defend your prior choice.



---

4. Reflect on the differences/similarities in the credible intervals between the two different priors as well as the frequentist model. If you were going to bet on the players shooting percentages for the next season, which would you prefer?


---

5. Now consider a hierarchical model where

$$FTM_i \sim Binomial(FTA_i, \theta_i)$$

$$\theta_i \sim  Beta(\omega*(\kappa-2)+1, (1-\omega)*(\kappa-2)+1)$$
$$\omega \sim Beta(2.55, .45) $$
$$\kappa-2 \sim Gamma(.01, .01)$$


Note that this specification of the Beta distribution uses a mean parameter $\omega$ and a dispersion parameter in $\kappa$.

Use the JAGS code to again estimate free throw shooting percentage and uncertainty intervals for the three players.

```{r}
# Model - see page 250 in text
modelString <- 'model {
  for ( s in 1:Nsubj ) {
    z[s] ~ dbin( theta[s], N[s] )
    theta[s] ~ dbeta( omega*(kappa-2)+1, (1-omega)*(kappa-2)+1)
  }
  omega ~ dbeta(2.55, .45) # 85%
  kappa <- kappaMinusTwo + 2
  kappaMinusTwo ~ dgamma(.01, .01)
}'
writeLines( modelString, con='HierModel.txt')
```

Also summarize your estimates for $\omega$ and $\kappa$


---

6. Finally, discuss the differences in all your estimation approaches. If you had to choose one analysis: HM, Bayes with uniform priors, Bayes with informative priors, or the frequentist approach which would you choose and why?