---
title: "Week 13 Activity: GLMs"
output: pdf_document
---

```{r setup, include=FALSE}
set.seed(04102023)
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rjags)
library(tidyverse)
```

  
Now reconsider the willow tit dataset and consider modeling not just the presence / absence of birds, but directly modeling the number of birds observed in each spatial region.



```{r, fig.align='center'}
birds <- read.csv('http://math.montana.edu/ahoegh/teaching/stat491/data/willowtit2013_count.csv')
head(birds) %>% kable()
```
\vfill

This dataset contains 242 sites and 6 variables:  

    - siteID, a unique identifier for the site, some were not sampled during this period 
    - elev, mean elevation of the quadrant in meters 
    - rlength, the length of the route walked by the birdwatcher, in kilometers 
    - forest, percent forest cover 
    - bird.count, number of birds identified 
    - searchDuration, time birdwatcher spent searching the site, in minutes



#### 1. Data Visualization
Create two figures that explore `bird.count` as a function of forest cover percentage (`forest`) and elevation (`elev`).

#### 2. Model Specification

Using a Poisson regression model, clearly write out the model to understand how forest cover and elevation impact bird count.
\vfill

#### 3. Priors
Describe and justify the necessary priors for this model.


\vfill

#### 4.  Fit MCMC
Fit the JAGS code for this model. 

#### 5. Summarize inferences from model
Talk about the model and discuss which and how predictor variables influence the observed bird count. Create figures that show your fitted model for elevation and/or forest coverage.


